GEE models

### 2017-10-15 doc/manuscript.Rmd
added all abbrev + rearrange order

beta cell function; diabetes pathogenesis; fatty acid composition insulin sensitivity; longitudinal cohort; non-esterified fatty acids;

@@

- GEE: Generalized estimating equations
- IGI/IR: Insulinogenic index over HOMA-IR
- ISI: Insulin Sensitivity Index (Matsuda index)
- ISSI-2: Insulin Secretion-Sensitivity Index-2
- ALT: Alanine aminotransferase
- TAG: Triacylglyceride
- PROMISE: Prospective Metabolism and Islet Cell Evaluation Cohort
- QIC: Quasi-Likelihood Information Criterion
- MAQ: Modified Activity Questionnaire
- WC: Waist circumference

@@

tab_basic <- tab('basic', 'Basic characteristics of the PROMISE participants at each of the three clinic visits.')
fig_dist <- fig('nefa', 'Concentrations (nmol/mL) of each NEFA in PROMISE participants at the baseline visit (2004-2006).')
fig_gee_unadj <- fig('gee_unadj', 'Longitudinal associations of individual non-esterified fatty acids (mol% and nmol/mL) with insulin sensitivity and beta cell function over 6 years in the PROMISE cohort. Models are only adjusted for time. X-axis values represent a percent difference in the outcome per SD increase in the fatty acid. P-values were adjusted for the BH false discovery rate, with each increase in dot size and blackness representing a p-value significance of p>0.05, p<0.05, p<0.01, and p<0.001.')
fig_gee_adj <- fig('gee_adj', 'Longitudinal associations of individual non-esterified fatty acids (mol% and nmol/mL) with insulin sensitivity and beta cell function over the 6 years in the PROMISE cohort. Generalized estimating equation models were fully adjusted. X-axis values represent a percent difference in the outcome per SD increase in the fatty acid. P-values were adjusted for the BH false discovery rate, with each increase in dot size and blackness representing a p-value significance of p>0.05, p<0.05, and p<0.01.')
```

@@

All analyses were performed using R `r getRversion()` [@Rbase], along with the R
packages geepack `r packageVersion('geepack')` for GEE [@Hoejsgaard2006a]
and the package pls `r packageVersion('pls')` for the PLS analysis.
The R code and output for all analyses of this manuscript is available at 
https://dx.doi.org/10.6084/m9.figshare.3472433. Results were considered
statistically significance at p<0.05, after adjusting for multiple testing using
the Benjamini-Hochberg False Discovery Rate [@Benjamini1995a]. STROBE was used
as a guideline for reporting [@Vandenbroucke2007b].
The cohort datasets analysed during the current study are available from the
corresponding author on reasonable request.

>>>>

beta cell function; diabetes pathogenesis; fatty acid composition; insulin sensitivity; longitudinal cohort; non-esterified fatty acids;

@@

- AR1: Autoregressive of order 1
- ALT: Alanine aminotransferase
- EPA: Eicosapentaenoic acid
- FDR: False Discovery Rate
- GEE: Generalized estimating equations
- HOMA: Homeostatic model assessment
- HOMA-IR: Homeostatic model assessment of insulin resistance
- HOMA2-%S: Homeostatic model assessment 2 of insulin sensitivity
- IFG: Impaired fasting glucose
- IGI/IR: Insulinogenic index over HOMA-IR
- IGT: Impaired glucose tolerance
- ISI: Insulin Sensitivity Index (Matsuda index)
- ISSI-2: Insulin Secretion-Sensitivity Index-2
- MAQ: Modified Activity Questionnaire
- MET: Metabolic Equivalent of Task
- NEFA: Non-esterified fatty acids
- OGTT: Oral glucose tolerance test
- PLS: Partial least squares regression
- PROMISE: Prospective Metabolism and Islet Cell Evaluation Cohort
- PUFA: Polyunsaturated fatty acids
- QIC: Quasi-Likelihood Information Criterion
- TAG: Triacylglyceride
- WC: Waist circumference

@@

tab_basic <- tab('basic', 'Basic characteristics of the PROMISE participants at each of the three clinic visits.')
fig_dist <- fig('nefa', 'Concentrations (nmol/mL) of each NEFA in PROMISE participants at the baseline visit (2004-2006).')
fig_gee_unadj <- fig('gee_unadj', 'Longitudinal associations of individual non-esterified fatty acids (mol% and nmol/mL) with insulin sensitivity and beta cell function over 6 years in the PROMISE cohort. Models are only adjusted for time. X-axis values represent a percent difference in the outcome per SD increase in the fatty acid. P-values were adjusted for the BH false discovery rate (FDR), with each increase in dot size and blackness representing a p-value significance of p>0.05, p<0.05, p<0.01, and p<0.001.')
fig_gee_adj <- fig('gee_adj', 'Longitudinal associations of individual non-esterified fatty acids (mol% and nmol/mL) with insulin sensitivity and beta cell function over the 6 years in the PROMISE cohort. Generalized estimating equation models were fully adjusted. X-axis values represent a percent difference in the outcome per SD increase in the fatty acid. P-values were adjusted for the BH false discovery rate (FDR), with each increase in dot size and blackness representing a p-value significance of p>0.05, p<0.05, and p<0.01.')
```

@@

All analyses were performed using R `r getRversion()` [@Rbase], along with the R
packages geepack `r packageVersion('geepack')` for GEE [@Hoejsgaard2006a]
and the package pls `r packageVersion('pls')` for the PLS analysis.
The R code and output for all analyses of this manuscript is available at 
https://dx.doi.org/10.6084/m9.figshare.3472433. Results were considered
statistically significance at p<0.05, after adjusting for multiple testing using
the Benjamini-Hochberg False Discovery Rate (FDR) [@Benjamini1995a]. STROBE was used
as a guideline for reporting [@Vandenbroucke2007b].
The cohort datasets analysed during the current study are available from the
corresponding author on reasonable request.
